\textit{A young boy with a tree made of trees}: \\ Domain Adaptation of an LSTM-based Image Caption Generation Model

### 2023-01-05 paper/paper.tex
Results (questionnaire)

\textit{Here we should summarize our results, in terms of:
\begin{enumerate}
    \item where did the performance exceed the best prior performance (we got BEST checkpoints?) \textbf{DK}
    \item which models actually generated reasonable captions? where did they start falling apart? \textbf{MS or DK}
    \item subjectively, which captions were best? include the questionnaire results here or subsequently \textbf{MS}
\end{enumerate}
Include figures and tables as much as we can/as much as is reasonable.}

@@

>>>>

\textit{Here we should summarize our results, in terms of:
\begin{enumerate}
    \item where did the performance exceed the best prior performance (we got BEST checkpoints?) \textbf{DK}
    \item which models actually generated reasonable captions? where did they start falling apart? \textbf{MS or DK}
\end{enumerate}
Include figures and tables as much as we can/as much as is reasonable.}

@@

While the captions themselves can be seen in the Jupyter Notebook files, the aim of the questionnaire was to reveal which model or models performed the best on which test set. Therefore, the results presented here are divided between the sets and include only the information on which model's caption was deemed the best, as well as how many people voted for it. Table 1 contains the information for the unfiltered imageCLEF set, Table 2 - for the filtered one, and Table 3 for Flickr8k. Image names are the file names without the .jpg extension. In some cases more than one model's caption received the same amount of votes, or the winning caption was generated by more than one model.

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{1cm}|p{4.5cm}|}
\hline \bf Image & \bf Best model (\% responses) \\ \hline
40416 & original (83.3\%)  \\
39158 & original (50\%)  \\
25053 & original (83.3\%)  \\
30620 & 100\_1.0 (66.7\%)  \\
32397 & 100\_0.2 (66.7\%) \\
38937 & original (50\%) \\
39005 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) or 100\_0.2 (33.3\%) \\
40120 & original (66.7\%) \\
23588 & 100\_1.0 (83.3\%)  \\
39472 & 100\_0.1 (66.7\%)  \\
30138 & original (50\%)  \\
40202 & 100\_0.1 (83.3\%)  \\
35895 & original (66.7\%)  \\
32663 & 100\_0.2 (83.3\%)   \\
30705 & original (83.3\%) \\
38081 & 100\_0.15 (66.7\%) \\
31571 & original (100\%) \\
35858 & original (66.7\%) \\
37836 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) \\
39239 & 100\_0.15, 100\_0.2, and 100\_1.0 (50\%)\footnote{These models generated identical captions.} \\
\hline
\end{tabular}
\end{center}
\caption{\label{filtered-table} Best captions per image in the filtered imageCLEF test set. }
\end{table}

As can be noticed in Table 1, the majority of the "good" captions for the unfiltered imageCLEF set were generated by the original, non-fine-tuned model, or by the one with a 10\% filter. Some captions from the model fine-tuned on nonfiltered data turned out to have been rated high as well. Finally, other models' captions were rated as good when they overlapped with the original model's caption. Overall, while the judgements are split rather evenly, the original model seems to generate better captions for this test set. It is worth pointing out that relatively few captions were unanimously voted best.

When it comes to the filtered imageCLEF test set (Table 2), where a maximum of 10\% tokens in the original caption could be unknown tokens, the spectrum of what models' captions were considered good is even wider. There was only one case where one caption was voted best by all the participants. Half of the best captions were generated by the original model, and half by various fine-tuned ones, with none of them being a clear front-runner in this case.

Finally, when it comes to the original Flickr8k-based test set, as seen in Table 3, the original model without fine-tuning performed best, although there were instances where the 10\% or 20\% fine-tuned models generated better captions. Many of the original model's captions were also chosen unanimously.

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2.5cm}|p{3cm}|}
\hline \bf Image & \bf Best model (\% responses) \\ \hline
2116444946[...] & original (100\%)  \\
2316097768[...] & original (100\%)  \\
2439384468[...] & original (66.7\%)  \\
2112921744[...] & 100\_0.2 (50\%)  \\
2392460773[...] & original (100\%) \\
2434006663[...] & original (100\%) \\
2308256827[...] & original (100\%) \\
2111360187[...] & original (66.7\%) \\
2271671533[...] & original (66.7\%)  \\
2328616978[...] & original (100\%)  \\
2456907314[...] & original (66.7\%)  \\
2229179070[...] & 100\_0.1 (100\%)  \\
2279980395[...] & 100\_0.1 (88.3\%)  \\
2393971707[...] & original (83.3\%)   \\
211277478[...] & 100\_0.1 (50\%) \\
2337919839[...] & original (83.3\%) \\
2447035752[...] & original (50\%) \\
23445819[...] & 100\_0.2 (66.7\%) \\
2448210587[...] & original (33.3\%) \\
2445654384[...] & 100\_0.1 (50\%) or 100\_0.2 (50\%)\\
\hline
\end{tabular}
\end{center}
\caption{\label{flickr-table} Best captions per image in the Flickr8k test set. }
\end{table}

When it comes to the comments left by the respondents after each section, they were not very positive. The captions in the first content section were judged as "unfitting," "strange," "ungrammatical," "inaccurate," "nonsensical," or "confusing," with some of the respondents admitting that the mismatch between the image and captions made them anxious or made their "brain hurt." For the second content section, some people considered the captions better, and some worse than previously, and the feeling of anxiety persisted here too. In addition, half of the respondents pointed out that the captions tend to include people in the descriptions even when there are no people in the images. It seems that the impressions of the captions in the last content section were better: while people said that they were still "strange," "weird," or "confusing," they also admitted that the verbs in the sentences tended to be accurate descriptions of the events in the images, and that they were overall "somewhat accurate." The person who expressed their feelings of anxiety previously now said that they "started to get used to them," which could also be a consequence of there being a smaller mis-match between the images and the captions.

In the final remarks many of the respondents expressed their dissatisfaction with the overall quality of the captions, leaving comments such as "r u b b i s h" or "my brain melted." Two people mentioned again how the captions would include people despite there not being any in the images. One person said that the simpler captions or images were clear, while the more complex ones were problematic or mismatched. Another person said that these "didn't sound like human-written captions," and yet another that they oftentimes "missed the essence of the picure."

Overall, the original model performed best on the original test set. The most captions generated by fine-tuned models were selected as best for the filtered imageCLEF test set. The unfiltered test set was also fairly balanced between the fine-tuned and original models. While for the unfiltered data the unfiltered and 10\% fine-tuned models can be identified as the best out of the fine-tuned ones, the same is impossible to determine for the filtered test set. From these results it can be concluded that while fine-tuning does improve the quality of the captions somewhat (both for imageCLEF and Flickr8k), it does not have that effect for every image. Judging by the open responses from the participants, the captions generated for Flickr8k were the least wrong, but their quality was still low, and the ones for imageCLEF were even worse.

Clearly, while some degree of improvement can be obtained by this kind of fine-tuning, it is far from perfect and the captions rarely match the images. This is likely due to the disparity between the kinds of images in the two sets. Another clue to that being a major issue is the models' tendency to - as pointed out by the respondents - insert people into captions when there were none in the original image. A large portion of the captions in the Flickr8k dataset are constructed that way, and it must be one of the things that the model learned to try to identify in the image at all cost and feature in the caption, and that the fine-tuning did not help mitigate.

Finally, it is worth keeping in mind that some of the apparent weirdness of the captions may stem from the relatively small vocabulary that the models had, which was likely lacking the proper words to describe certain elements in the images, even when the ones whose captions contained many new words were filtered out.
